---
title: "AugustineUgbeda_R PROJECT/HOTEL PREDICTION"
author: "AUGUSTINE UGBEDA"
date: "2020 M11 18"
output:
  html_document: default
  pdf_document: default
---


RUN NESSASARY LIBRARIES

```{r}
library(tidyverse) #for data cleaning
library(Boruta) #for feature selection
library(randomForest)
library(lattice)
library(mlbench)
library(dplyr)
library(tidyselect)
library(caret)  # package for model training(Regression and classifications)
library(ggplot2) # package for creating charts/visualizations
library(e1071) #needed for random forest and confusion matrix
```
IMPORTING DATASET and DATA PREPARATION
```{r}
HOTEL <- read.csv("C:\\Users\\augustine ugbeda\\Documents\\R\\R PROJECT\\H1.csv",stringsAsFactors = T)

```
REMOVAL OF NAs
```{r}
hotel <- na.omit(HOTEL)
```

```{r}
###RANDOM FOREST cannot handle variables with over 53 categories ,therefore i need to check the columns with over 53 categories and drop them 
str(HOTEL)

```

```{r}
variables <- c("Company","ReservationStatusDate","Agent","Country","ReservationStatus") # list of variabees
hotel <- hotel[ , !(names(hotel)%in% variables)] #dropping the variables

```

DECLARING the independent variable(IsCanceled) as a factor
```{r}
# declearing IsCanceled as a factor
hotel$IsCanceled<-as.factor(hotel$IsCanceled)
```




SUMMARY STATISTICS
```{r}
summary(hotel)

```
STRUCTURE
```{r}
str(hotel)

```





###VISUALIZATIONS
```{r}
#boxplot visualisations of IsCancelled against other variables
par(mfrow=c(3,3))

boxplot(LeadTime~IsCanceled,ylab = "LeadTime",xlab="IsCanceled",col="blue",data= hotel)
boxplot( ArrivalDateYear ~IsCanceled,ylab = "ArrivalDateYear",xlab="IsCanceled",col="yellow",data= hotel)
boxplot(ArrivalDateWeekNumber ~IsCanceled,ylab = "ArrivalDateWeekNumber",xlab="IsCanceled",col="green",data= hotel)
boxplot(Children ~IsCanceled,ylab = "Children",xlab="IsCanceled",col="light blue",data= hotel)
boxplot(Babies~IsCanceled,ylab = "Babies",xlab="IsCanceled",col="orange",data= hotel)

boxplot(PreviousCancellations ~IsCanceled,ylab = "PreviousCancellations",xlab="IsCanceled",col="violet",data= hotel)
boxplot(ArrivalDateDayOfMonth ~IsCanceled,ylab = "ArrivalDateDayOfMonth",xlab="IsCanceled",col="violet",data= hotel)
```


```{r}
xyplot(LeadTime~ArrivalDateDayOfMonth|IsCanceled,
       layout=c(3,1),#with ncol=3, and nrow=1
       group=IsCanceled,data = hotel,
       type=c("p","smooth"),
       scales = "free")#show points on smooted line
```
```{r}
ggplot(hotel,aes(x=ArrivalDateMonth, y=DaysInWaitingList,group=1)) + stat_summary(fun="mean", geom="line", col="green") + 
  ggtitle("Average Days on Waiting List by Arrival Month") + ylab("Average Days on WaitingList") + xlab("Month") +theme_bw() 
```
The line graph above shows that december has the highest average days in waiting list




CREATING THE TEST AND TRAINING DATA
```{r}
#create training set(SPLITING THE TARGET VARIABLE)
table(hotel$IsCanceled)
data1 <-hotel[which(hotel$IsCanceled==0),] 
data2 <-hotel[which(hotel$IsCanceled==1),]

set.seed(100)

#this shows that we have 28938 for not cancelled and 11122 for cancelled
```

```{r}
##set training data to 80% of the dataset
training1 <- sample(1:nrow(data1),0.80*nrow(data1))
training2 <-sample(1:nrow(data2),0.80*nrow(data2))

View(training1)
View(training2)

```

```{r}
#put both in data frames
training_1 <- data1[training1,]
View(training_1)
training_2 <- data2[training2,]
View(training_2)

```

```{r}
trainingdata <- rbind(training_1,training_2)##combine the two
View(trainingdata)
```

```{r}
##create test data 20 %
test_1 <- data1[-training1,]#contain 20 %of those not canceled
test_2 <- data2[-training2,]#contains 20%of those canceled

testdata <- rbind(test_1,test_2)
View(testdata)
```


FEATURE SELECTION USING RANDOM FOREST


```{r}
set.seed(150)
model <- randomForest(IsCanceled~.,data = hotel)
RandomF <- importance(model)
RandomF
```

```{r}
# Converting the result got into a data frame for visualizations
RandomF_DF <- data.frame(Variables = row.names(RandomF), MeanDecreaseGini = RandomF[,1])
RandomF_DF <- RandomF_DF[order(RandomF_DF$MeanDecreaseGini, decreasing = TRUE),]
RandomF_DF
```


```{r}

# Graphical represtation of the feature selection
ggplot(RandomF_DF[1:20,], aes(x=reorder(Variables, MeanDecreaseGini), y=MeanDecreaseGini, fill=MeanDecreaseGini)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% of increase in MeanDecreaseGini if variable is randomly permuted') + coord_flip() + theme(legend.position="none")

```
from the above visualization,we can see that "leadtime" has more effect than all the other variables in the dataset

for the purpose of creating the model,we will use variables with MeanDecreaseGini of 500 and above


MODEL BUILDING
```{r}
##building the logistic regression models
logisticmodel1 <- glm(IsCanceled~LeadTime+ADR+DepositType+MarketSegment+RequiredCarParkingSpaces+StaysInWeekNights+CustomerType, data = trainingdata, family = binomial(link = "logit"))

summary(logisticmodel1)



```


predict scores and add to the data
```{r}
testdata$predict<- predict(logisticmodel1,testdata)
View(testdata)
```

```{r}
AIC(logisticmodel1) #penalty for additional variable
BIC(logisticmodel1)# a stronger penalty for adding additional variable
```
```{r}
# Model1 Accuracy
testdata$predict_1 <- ifelse(testdata$predict >0, 1,0)
View(testdata)
```



checking if the predict_1 column is factor  and changing the predict_1 column to factor 
```{r}
is.factor(testdata$predict_1)

testdata$predict_1<- as.factor(testdata$predict_1)

testdata$IsCanceled<- as.factor(testdata$IsCanceled)

```


CONFUSION MATRIX
```{r}
confusionMatrix(testdata$IsCanceled, testdata$predict_1)
```



```{r}
#PRACTICAL REAL TIME PREDICTION
#Lets say a person books for a reservation and have a profile as below
#now you want to predict the chances that the person will cancel his reservation
guest<- data.frame(DepositType = "No Deposit     ", ADR = 0.00, MarketSegment = "Corporate", RequiredCarParkingSpaces = 1, StaysInWeekNights = 2,CustomerType = "Contract",LeadTime = 200)

predictions<- predict(logisticmodel1, guest)
predictions
guest$pred <- ifelse(predictions>0, "1","0")
guest



```

from the above we can see the predicted value for the guest is 0 which is Not canceled.therefore we predict that a guest with the above profile will not cancel his or her reservation